MetaTracts - A Method for Robust Extraction and Visualization of Carbon Fiber Bundles in Fiber Reinforced Composites

### 2014-09-04 chapters/MetaTracts.tex
updated image,

\begin{figure}
\includegraphics[width=0.5\textwidth,clip=true, trim= 0cm 0cm 0cm 0cm]{imagesMT2014/MT_data_4_2.png}
\caption{Data Characteristics. A: Shows a rendering of the data. B: Shows 2D slices along the X (Right) and Z (Left) axis. C: Shows a zoom of the yellow square, multiple bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}

@@

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{imagesMT2014/algo_A.png}
\caption{Flow chart of our approach}
\label{fig:flowchart}
\end{figure}
 Figure \ref{fig:flowchart} shows our pipeline. The approach consists of multiple steps which we divide into two broad phases, extracting the MetaTracts  (this section) and  finding plausible fiber bundles (Section~\ref{subsec:fiber-bundles}).
%\subsection {Extracting MetaTracts}
%\label{subsec:ExFiberCyl}
%We explain the different steps of MetaTracts extraction.
\subsection {Reliable Hessians}
\label{subsec:rh}

@@

\begin{figure}[h]
  \begin{center}
  \includegraphics[ width=0.9\linewidth]{imagesMT2014/MT_abstract_A2} %algo-1d}
  \end{center}
 \caption{MetaTracts generation. B. MetaTract generation, region with $R_H$ less than threshold. A: close up of marked region in A. C: Parts of a single MetaTract.}
  \label{fig:algo}

@@

 In this particular example Figure~\ref{fig:algo}B we select the orange grid vertex next and repeat the process. We do not select the purple vertex (Figure~\ref{fig:algo}C) because it is not a candidate vertex. If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and will be of short length. Short tracts are then remove. The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}(a). The MetaTracts are colored with the mean orientation direction mapped to the RGB space. That consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
 We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.

@@

\begin{figure}[h]
\centering
%	\subfloat[]{\includegraphics[width=0.49\linewidth]{imagesMT2014/MT_crop16_MT_2.png}}
	\subfloat[]{\includegraphics[width=0.49\linewidth]{imagesMT2014/MT_crop16_MT_2.PNG}} 
	\subfloat[]{\includegraphics[width=0.49\linewidth]{imagesMT2014/length_distribution.PNG}}\\
%	\subfloat[]{\includegraphics[width=0.9\linewidth]{imagesMT2014/mt_crop_16_orient.PNG}}
	\caption{(a) MetaTracts colored according to mean orientations. (b) length distribution of individual MetaTracts for a particular bundle.}
	\label{fig:length_distribution}
\end{figure}

@@

However, the MetaTracts produced as described in Section 4 often do not extend the full length of the fiber bundle.  Thus, MetaTracts
in a given fiber bundle may have very different lengths and may only partially overlap Figure~\ref{fig:length_distribution}(b).

@@

  \begin{figure}
  \centering
  	%\includegraphics[ trim = 2cm 2cm 2cm 2cm, clip=true, width=0.25\textwidth]{imagesMT2014/orientation_clustering_A.png}~
  	%	\includegraphics[ trim = 2cm 2cm 2cm 2cm, clip=true, width=0.25\textwidth]{imagesMT2014/orientation_clustering_A.png}
%  	\captionsetup[subfigure]{labelformat=empty}
%  	\subfloat[1]{\includegraphics[width=0.2\linewidth]{imagesMT2014/orientation_clustering_A.png}}
%  	\subfloat[2]{\includegraphics[width=0.2\linewidth]{imagesMT2014/crop_16_clustering_1.PNG}}
% 	\subfloat[3]{\includegraphics[width=0.2\linewidth]{imagesMT2014/crop_16_clustering_2.PNG}}
% 	\subfloat[4]{\includegraphics[width=0.2\linewidth]{imagesMT2014/curveD.PNG}}
%  	
 \includegraphics[width=0.5\textwidth]{imagesMT2014/crop-16/Capture4.PNG}
  	\caption{(B)Shows the result of K means clustering with data points projected to the top three eigenvectors as major axes. (A,C) show the MetaTracts colored according to the clustering results. (The white to black tracts are only to show context.)  }
  \label{fig:orientation_clustering}
  \end{figure}

@@

 The Hausdorff distance is a metric so $H(P,Q) \le H(P,Q') + H(Q',Q)$
 but the directed Hausdorff is not.
 Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTract which only partially cover the bundle (figure~\ref{fig:length_distribution}(b)).
 If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$,
 then $H_{dir}(P,Q)$ will be very small while $H_{dir}(Q,P)$ will be large.
 Thus, $H(P,Q)$ will be large, even though $P$ and $Q$ are in the same
 fiber bundle.

@@

The second is to extract the corresponding surfaces from the voxel data by binarizing the volume per cluster and extracting the isosurface of the largest connected component from the binary volume. To voxelize the space,  we compute a neighborhood around each voxel. We then enumerate the number of voxels of each class (cluster) in this neighborhood. The voxel is then assigned to the class with the maximum number of elements in the neighborhood and Fig.~\ref{fig:mesh}A shows the result of voxelization. Fig.~\ref{fig:mesh}(B, C) show examples of extracted meshes.

@@

\section {Experimental Results}
\label {sec:results}

%\begin{figure*}
%\centering
%\captionsetup[subfigure]{labelformat=empty}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-G.PNG}}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-H.PNG}}
%	%\subfloat[]{\includegraphics[width=0.35\textwidth]{imagesMT2014/crop-6-I.PNG}}
%	
%	\subfloat[(a) Dat set 2]{\includegraphics[width=0.7\linewidth]{imagesMT2014/prepreg/Capture8.PNG}}
%	\subfloat[(b) Data set 3]{\includegraphics[width=0.2\linewidth]{imagesMT2014/dataset203.png}}	
%	\caption{(a) Shows in order, the volume rendering of the second data set, a single (2D) slice, the orientation clustering bundles and the final bundles followed by proximity clustering. (b) Shows in order the third data set, a single (2D) slice and finally the extracted fiber bundles.}
%	\label{fig:crop-6}
%\end{figure*}

\begin{figure}[h]
\centering
	\includegraphics[width=0.5\textwidth]{imagesMT2014/all_clusters.PNG}
	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.}
	\label{fig:crop-16-decomp}
\end{figure}
\begin{figure}[h]
  \centering
   \includegraphics[width=0.5\textwidth]{imagesMT2014/mesh_2_C.png}
   \caption{Volumes and  meshes. (A) Shows the voxelized volume. (B) Shows two different extracted meshes (C) shows the combined meshes.}
   \label{fig:mesh}
   \end{figure}
\begin{figure}
\centering
	\includegraphics[width=0.5\textwidth]{imagesMT2014/crop-6-J_2.PNG}
	\caption{Data set with flat thin and compact bundles. A shows the volume rendering and a 2D slice with one of the boundaries marked in red, B shows the clusters according to individual orientation.C shows the complete result. D shows the voxelization of C.}
	\label{fig:prepreg}
\end{figure}
\begin{figure}
\centering
	\includegraphics[width=0.5\textwidth]{imagesMT2014/Graph_crop16.PNG}
	%\includegraphics[width=0.23\textwidth]{imagesMT2014/Graph_crop16_3.PNG}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}C, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top). The unit for length is the grid cube length. Figure~\ref{fig:length_distribution}(b) shows the length of individual MetaTracts in cluster 6.}
	\label{fig:len_dist_crop16}
\end{figure}

@@

 %glass
We tested our method on multiple datasets with varying characteristics. 
Figure~\ref{fig:crop-16-decomp} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into 10 clusters (the ground truth is 6 and 5 respectively). The hierarchical clustering is robust to over segmentation.
Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster. We generate the correct clusters and outlier bundles with very few elements which can be discarded. As noted in section~\ref{subsec:dist_clustering} the tracts donot run the entire length of the dataset, the median length of MetaTracts in cluster 6 is 200 while the maximum MetaTract in the bundle is about 500 (measured in units of grid cube length). Fig.~\ref{fig:length_distribution}(b) shows the length per MetaTract distribution for a particular fiber bundle.

@@

\section {User Evaluation}
\label {sec:evaluation}
\begin{figure}[h]
\centering
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/crop-6-F.PNG}}
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/orientB.PNG}}
\caption{Contextual information: (a) Shows the four largest clusters in the data set in Fig~\ref{fig:prepreg}.
(b)Shows how the red bundle `bends' under the green bundle, which is not immediately apparent from just the volume rendering.}
\label{fig:context}
\end{figure}

\section{Parameter choices}
\label{sec:param_choices}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions. Which is generally known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$ (Section~\ref{sec:results}). Large $h$ keeps the major bundles intact. For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4 and 5 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements). This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.

>>>>

\begin{figure}
\includegraphics[width=0.5\textwidth,clip=true, trim= 0cm 0cm 0cm 0cm]{imagesMT2014/MT_data_4_2.png}
\caption{Data Characteristics. A: Shows a rendering of the data. B: Shows 2D slices along the X (Right) and Z (Left) axis. C: Shows a zoom of the green square, multiple bundles cross and are indistinguishable by visual inspection alone. }
\label{fig:data-char}
\end{figure}
\section {Data Characteristic and Assumptions}
\label{sec:char_data}

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{imagesMT2014/algo.png}
\caption{Flow chart of our approach}
\label{fig:flowchart}
\end{figure}
 Figure \ref{fig:flowchart} shows our pipeline. The approach consists of multiple steps which we divide into two broad phases, extracting the MetaTracts  (this section) and  finding plausible fiber bundles (Section~\ref{subsec:fiber-bundles}).
%\subsection {Extracting MetaTracts}
%\label{subsec:ExFiberCyl}
%We explain the different steps of MetaTracts extraction.
\subsection {Reliable Hessians}
\label{subsec:rh}

@@

\begin{figure}[h]
  \begin{center}
  \includegraphics[ width=0.4\textwidth]{imagesMT2014/MT_abstract} 
  \end{center}
 \caption{MetaTracts generation. B. MetaTract generation, region with $R_H$ less than threshold. A: close up of marked region in A. C: Parts of a single MetaTract.}
  \label{fig:algo}

@@

 In this particular example Figure~\ref{fig:algo}B we select the orange grid vertex next and repeat the process. We do not select the purple vertex (Figure~\ref{fig:algo}C) because it is not a candidate vertex. If we generate tracts that have erroneous local orientations they will not be able to find further possible candidate vertices and will be of short length. Short tracts are then remove. The MetaTracts generated are shown in Figure~\ref{fig:length_distribution}(A). The MetaTracts are colored with the mean orientation direction mapped to the RGB space. That consistent orientation is a key intrinsic feature in the data becomes visually pronounced.
 We apply uniform, dense seeding to the XCT volume data to trace and generate the tracts of fiber bundles.

@@

\begin{figure}[h]
\centering
	\includegraphics[width=0.4\textwidth]{imagesMT2014/crop-16/length_dist.png}
	\caption{(A) MetaTracts colored according to mean orientations. (B) length distribution of individual MetaTracts for a particular bundle.}
	\label{fig:length_distribution}
\end{figure}

@@

However, the MetaTracts produced as described in Section 4 often do not extend the full length of the fiber bundle.  Thus, MetaTracts
in a given fiber bundle may have very different lengths and may only partially overlap Figure~\ref{fig:length_distribution}(B).

@@

  \begin{figure}
  \centering
  	%\includegraphics[ trim = 2cm 2cm 2cm 2cm, clip=true, width=0.25\textwidth]{imagesMT2014/orientation_clustering_A.png}~
  	%	\includegraphics[ trim = 2cm 2cm 2cm 2cm, clip=true, width=0.25\textwidth]{imagesMT2014/orientation_clustering_A.png}
%  	\captionsetup[subfigure]{labelformat=empty}
%  	\subfloat[1]{\includegraphics[width=0.2\linewidth]{imagesMT2014/orientation_clustering_A.png}}
%  	\subfloat[2]{\includegraphics[width=0.2\linewidth]{imagesMT2014/crop_16_clustering_1.PNG}}
% 	\subfloat[3]{\includegraphics[width=0.2\linewidth]{imagesMT2014/crop_16_clustering_2.PNG}}
% 	\subfloat[4]{\includegraphics[width=0.2\linewidth]{imagesMT2014/curveD.PNG}}
%  	
 \includegraphics[width=0.5\textwidth]{imagesMT2014/crop-16/orientation_cluster.PNG}
  	\caption{(C) Shows the result of K means clustering with data points projected to the top three eigenvectors as major axes. (A) shows the MetaTracts projected on top three eigenvectors. (B) shows 2D slices of C. }
  \label{fig:orientation_clustering}
  \end{figure}

@@

 The Hausdorff distance is a metric so $H(P,Q) \le H(P,Q') + H(Q',Q)$
 but the directed Hausdorff is not.
 Unfortunately, the Hausdorff distance does not work well for our application since a fiber bundle may have many MetaTract which only partially cover the bundle (figure~\ref{fig:length_distribution}(B)).
 If a MetaTract $P$ covers only part of the fiber bundle covered by $Q$,
 then $H_{dir}(P,Q)$ will be very small while $H_{dir}(Q,P)$ will be large.
 Thus, $H(P,Q)$ will be large, even though $P$ and $Q$ are in the same
 fiber bundle.

@@

The second is to extract the corresponding surfaces from the voxel data by binarizing the volume per cluster and extracting the isosurface of the largest connected component from the binary volume. To voxelize the space,  we compute a neighborhood around each voxel. We then enumerate the number of voxels of each class (cluster) in this neighborhood. The voxel is then assigned to the class with the maximum number of elements in the neighborhood and Fig.~\ref{fig:crop-16-decomp}C and Fig~\ref{fig:prepreg}D shows the result of voxelization. Fig.~\ref{fig:crop-16-decomp}(E, F) show examples of extracted meshes.

@@

\section {Experimental Results}
\label {sec:results}
%\begin{figure*}
%\centering
%\captionsetup[subfigure]{labelformat=empty}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-G.PNG}}
%	%\subfloat[]{\includegraphics[width=0.10\textwidth]{imagesMT2014/crop-6-H.PNG}}
%	%\subfloat[]{\includegraphics[width=0.35\textwidth]{imagesMT2014/crop-6-I.PNG}}
%	
%	\subfloat[(a) Dat set 2]{\includegraphics[width=0.7\linewidth]{imagesMT2014/prepreg/Capture8.PNG}}
%	\subfloat[(b) Data set 3]{\includegraphics[width=0.2\linewidth]{imagesMT2014/dataset203.png}}	
%	\caption{(a) Shows in order, the volume rendering of the second data set, a single (2D) slice, the orientation clustering bundles and the final bundles followed by proximity clustering. (b) Shows in order the third data set, a single (2D) slice and finally the extracted fiber bundles.}
%	\label{fig:crop-6}
%\end{figure*}
\begin{figure*}[t]
\centering
	\includegraphics[width=0.45\textwidth]{imagesMT2014/crop-16/bundles}~ 
	\includegraphics[height=0.35\textwidth]{imagesMT2014/crop-16/vol_mesh.png} 
	\caption{(A) Orientation wise decomposition of data set Figure~\ref{fig:data-char}A into the extracted bundles and (B) all of them combined.(C) Shows the results of voxelization of B. (D) shows a 2D slice of the volume. (E,F) shows the extracted meshes.}
	\label{fig:crop-16-decomp}
\end{figure*}
\begin{figure}
\centering
	\includegraphics[width=0.35\textwidth]{imagesMT2014/crop-6.PNG}
	\caption{Data set with flat thin and compact bundles. A shows the volume rendering and a 2D slice with one of the boundaries marked in red, B shows the clusters according to individual orientation.C shows the complete result. D shows the voxelization of C.}
	\label{fig:prepreg}
\end{figure}

@@

 %glass
We tested our method on multiple datasets with varying characteristics. 
Figure~\ref{fig:crop-16-decomp} shows the result of data set~\ref{fig:data-char} decomposed into 2 orientation and each orientation decomposed into 10 clusters (the ground truth is 6 and 5 respectively). The hierarchical clustering is robust to over segmentation.
Figure~\ref{fig:len_dist_crop16} shows the median, minimum and maximum lengths per cluster. We generate the correct clusters and outlier bundles with very few elements which can be discarded. As noted in section~\ref{subsec:dist_clustering} the tracts donot run the entire length of the dataset, the median length of MetaTracts in cluster 6 is 200 while the maximum MetaTract in the bundle is about 500 (measured in units of grid cube length). Fig.~\ref{fig:length_distribution}(B) shows the length per MetaTract distribution for a particular fiber bundle.

@@

\section {User Evaluation}
\label {sec:evaluation}
\begin{figure}[h]
\centering
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/crop-6-F.PNG}}
\subfloat[]{\includegraphics[width=0.25\textwidth]{imagesMT2014/orientB.PNG}}
\caption{Contextual information: (a) Shows the four largest clusters in the data set in Fig~\ref{fig:prepreg}.
(b)Shows how the red bundle `bends' under the green bundle, which is not immediately apparent from just the volume rendering.}
\label{fig:context}
\end{figure}
\section{Parameter choices}
\label{sec:param_choices}
\begin{figure}
\centering
	\includegraphics[width=0.5\textwidth]{imagesMT2014/crop-16/histogram_10_A.PNG}
	\caption{Number of tracts and median, minimum and maximum length of individual MetaTracts in the orientation cluster Fig.~\ref{fig:orientation_clustering}, clustered into 10 clusters (Fig.~\ref{fig:crop-16-decomp}A top). The unit for length is the grid cube length. Figure~\ref{fig:length_distribution}(B) shows the length of individual MetaTracts in cluster 6.}
	\label{fig:len_dist_crop16}
\end{figure}
The critical parameters are $K$ for the K-means in orientation based clustering and $h$ for the hierarchical clustering. K denotes the number of major directions. Which is generally known a priori or can be easily estimated by looking at the weaving pattern. 
Our framework is robust to the choice of $h$ (Section~\ref{sec:results}). Large $h$ keeps the major bundles intact. For example, in  Fig.~\ref{fig:len_dist_crop16}, where $h$ was set to ten while the ground truth was five, we observe that the major clusters remain well segmented and small clusters can be removed if they have too few elements (clusters 2,3,4 and 5 in Fig.~\ref{fig:len_dist_crop16} all have less than 10 elements). This is an appealing trait of the proximity based clustering, thus providing good results even when exact $h$ might be unknown.
